---
title: "peerassessment1"
output: html_document
---

The code below loads the activity data and preprocesses it by removing observations with unknown steps ("NA").


```{r, echo = TRUE}
activitydata <- read.csv("activity.csv")
activitydataclean <- activitydata[!is.na(activitydata$steps),]
```

The total number of steps for each day are calculated and plotted using a histogram
```{r, echo = TRUE}
totalsteps <- aggregate(steps ~ date, data=activitydataclean, FUN = function(x) c(S = sum(x)))
hist(totalsteps$steps)
```
The mean of total steps is
```{r,echo=TRUE}
mean(totalsteps$steps)
```

The median steps are

```{r, echo = TRUE}
median(totalsteps$steps)
```

The averate daily activity pattern looks like below
```{r, echo = TRUE}
avgintervalsteps <- aggregate(steps ~ interval, data=activitydataclean, FUN = function(x) c(M = mean(x)))
plot(avgintervalsteps$interval,avgintervalsteps$steps,type = "l")
```

The maximum number of average steps for a given 5 minute interval is

```{r, echo = TRUE}
max(avgintervalsteps$steps)
```

the maximum is in the 5 minute interval ending with

```{r, echo = TRUE}
avgintervalsteps[avgintervalsteps$steps == max(avgintervalsteps$steps), 1]
```

The total number of missing values in the dataset are 
```{r, echo = TRUE}
sum(is.na(activitydata$steps))
```

The missing data is filled with average number os steps in 5 minute interval
```{r, echo = TRUE}
activitydatafilled <- activitydata
activitydatafilled$steps[is.na(activitydata$steps)]=mean(activitydataclean$steps)

```

The histogram with the new data set with filled in missign values looks like below

```{r, echo = TRUE}
totalstepsfilled <- aggregate(steps ~ date, data=activitydatafilled, FUN = function(x) c(S = sum(x), A= mean(x), M = median(x)))
hist(totalstepsfilled$steps)
```

The mean and median steps per day with the new data set are

The mean of total steps is
```{r,echo=TRUE}
mean(totalstepsfilled$steps)
```

The median steps are

```{r, echo = TRUE}
median(totalstepsfilled$steps)
```

and are very different to the ones before. 


Now this data set is added a factor variable "weekday" and "weekend"

```{r, echo = TRUE}
activitydatafilledweekday <- activitydatafilled
activitydatafilledweekday$weekend  = chron::is.weekend(activitydatafilledweekday$date)
activitydatafilledweekday$weekend[activitydatafilledweekday$weekend] <- "weekend"
activitydatafilledweekday$weekend[activitydatafilledweekday$weekend == "FALSE"] <- "weekday"
```

A panel plot of steps during weekend and weekdays with the filled-in dataset looks like below

```{r, echo = TRUE}
activitydatafilledweekdaytrue <- activitydatafilledweekday[activitydatafilledweekday$weekend == "weekday",]
activitydatafilledweekdayfalse <- activitydatafilledweekday[activitydatafilledweekday$weekend == "weekend",]

avgintervalstepsweekday <- aggregate(steps ~ interval, data=activitydatafilledweekdaytrue, FUN = function(x) c(M = mean(x)))
avgintervalstepsweekend <- aggregate(steps ~ interval, data=activitydatafilledweekdayfalse, FUN = function(x) c(M = mean(x)))

library(ggplot2)
p1 <- ggplot(avgintervalstepsweekday,aes(x = interval, y = steps)) + geom_line() + ggtitle("Weekday")
p2 <- ggplot(avgintervalstepsweekend,aes(x = interval, y = steps)) + geom_line() + ggtitle("Weekend")

library(gridExtra)
grid.arrange(p1,p2, ncol = 1)

```



